---
title: "PROJEC1-REPORT"
author: "Vo_Phi_Son, Nguyen Minh Tuan, Vo Diep Nhu"
date: "2024-03-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# INTRODUCTION

TThe dataset originating from the Mr. Trash Wheel Baltimore Healthy Harbor initiative embodies a significant real-world impact in combating environmental degradation by removing trash from waterways. This initiative is pivotal in addressing pressing environmental concerns, emphasizing the paramount importance of environmental conservation. However, working with this dataset may pose certain challenges, such as data inconsistencies or gaps, necessitating meticulous preprocessing and analytical methodologies to derive meaningful insights accurately. Within this dataset lie multifaceted dimensions of data, including trash weight, volume, and specific item counts like plastic bottles and cigarette butts. Such diverse data attributes offer a rich landscape for exploration and visualization, enabling comprehensive analyses of trash accumulation and composition. Through techniques such as line charts, bar charts, and box plots, temporal and comparative analyses become feasible, facilitating the identification of trends, seasonal variations, and disparities among different Trash Wheels. Moreover, the dataset's unique attribute, "HomesPowered," underscores the potential renewable energy benefits of waste management efforts, thereby reinforcing the significance of recycling and resource recovery in environmental sustainability endeavors. Overall, this dataset presents an invaluable opportunity to delve into environmental issues through the lens of data visualization, fostering a deeper understanding of the intricate dynamics between human activity and environmental well-being. Our group employs two questions to analyze the data, alongside various charts including bar charts, line charts, and box plots, to interpret the significance of the data. 
# Question 1: 
How does the composition of collected trash vary across different Trash Wheels (ID) over time (Month and Year)? 
## Introduction: 
Understanding the composition of collected trash across different Trash Wheels over time (Month and Year) is pivotal for comprehensively addressing environmental concerns. This analysis aims to explore the relationship between the quantities of plastic bottles, polystyrene items, and glass bottles collected by each Trash Wheel in various months and years. By delving into this inquiry, we seek to uncover patterns and variations in waste composition, which can offer crucial insights into the effectiveness of waste management strategies and environmental conservation efforts. To address this question, essential components of the dataset include information regarding the Trash Wheel identifiers (ID), the quantities of plastic bottles, polystyrene items, and glass bottles collected, and the temporal attributes of Month and Year. These elements enable us to track the distribution and fluctuations of different types of waste across different Trash Wheels over time.

This question is of significant interest due to its direct relevance to environmental sustainability and waste management practices. Understanding how the composition of collected trash varies across different Trash Wheels provides valuable insights into the sources and distribution of pollutants in water bodies. Moreover, uncovering trends and patterns in waste composition can inform targeted interventions and policy initiatives aimed at mitigating pollution and promoting ecosystem health. Therefore, this analysis holds the potential to contribute meaningfully to the ongoing efforts towards environmental conservation and sustainable waste management. 
## Approach: 
To address our question, we will employ a variety of plot types to glean insights into the composition and trends of collected waste over time. Initially, we will undertake two steps. In the first step, we will preprocess the data, creating new columns tailored to directly address our inquiry. Subsequently, in the second step, we will utilize line charts and bar charts to elucidate the temporal trends of waste types by month and year.

Firstly, we will employ a line chart to illustrate the temporal trends of weight and volume of collected waste across different years. The line chart is an ideal choice for showcasing continuous trends over time, allowing us to observe fluctuations and patterns in waste quantities throughout the years. By visually representing the changes in weight and volume over time, we can discern any significant peaks, valleys, or overall trends in waste accumulation.

Furthermore, we will create additional plots such as stacked area plots to compare the composition of collected trash among different Trash Wheels by months. Stacked area plots effectively convey the cumulative composition of waste types over time, enabling easy comparison of proportions among different categories within each month. Additionally, to provide a comprehensive understanding of waste distribution, we will employ a relative frequency bar chart encoded with color mapping. This chart will display the percentage of each waste type over the months, allowing us to identify any shifts in waste composition over time.

Lastly, we will investigate the trend of weight and volume of collected waste specifically in the year 2018. For this purpose, we will utilize a combination of line chart and bar chart. This approach will provide a holistic view of waste trends in 2018, leveraging the strengths of both chart types to depict variations in weight and volume across different months. By employing a diverse range of plot types, we aim to uncover nuanced insights into the dynamics of collected waste over time, facilitating informed decision-making and environmental management strategies. 

## Analysis: 
To answer the question 1, we do 2 steps. 
+ Step 1: We clear the data and create new columns to directly answer the question. 
+ Step 2: We plot the line chart and bar chart to describe the trend of types of waste by month and year.

```{r}
library(tidyverse)
```

We load the data from the github

```{r}
# TWCD_MrTrashWheel <- read_csv("/Users/tuan/Desktop/vinuni-stuff/comp5120-datavis/projects/Trash Wheel Collection Data.xlsx - Mr. Trash Wheel.csv")
trashwheel <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-03-05/trashwheel.csv')
```

Some data at the HomesPowered column is missed so we add the missing information by using the formulate: $HomePowered=\frac{Weight\ *\ 500}{30}$

```{r}
# change the value of HomesPowered = Weight * 500 / 30
trashwheel <- trashwheel %>% 
    mutate(HomesPowered = round(Weight * 500 / 30, 0.5))
```

We admit the null data

```{r}
trashwheel <- na.omit(trashwheel)
```

We want to investigate the trend of weight (in tons) and volume (in cubic yards) of wastes which were collected during the year from 2014 to 2022. Then, we sum the weight and volume of the wastes with grouping by year.

```{r}
sum_trashwheel <- trashwheel %>% 
    group_by(Year) %>% 
    summarise(sum_Weight = sum(Weight), sum_Volume = sum(Volume), sum_PlasticBottles = sum(PlasticBottles), sum_Polystyrene = sum(Polystyrene), sum_CigaretteButts = sum(CigaretteButts), sum_GlassBottles = sum(GlassBottles), sum_PlasticBags = sum(PlasticBags), sum_Wrappers = sum(Wrappers), sum_SportsBalls = sum(SportsBalls), sum_HomesPowered = sum(HomesPowered))
```

After that, we plot the weight and voulume of collected waste in year.

```{r}
# show the Volume and Weight by year
ggplot(sum_trashwheel,
    aes(x = Year, y = sum_Weight)) + 
    geom_line(aes(color = "Weight")) + 
    geom_line(aes(x = Year, y = sum_Volume, color = "Volume")) + 
    labs(title = "Fig 1. Weight and Volume of collected waste by Year", x = "Year", y = "Weight (tons) and Volume (cubic yards)", color = "Type") + 
    scale_color_manual(values = c("Weight" = "red", "Volume" = "blue")) +
    theme_minimal() + theme(
  plot.title = element_text(color = "red", face = "bold", hjust = 0.5))
```

### Discussion:

The Fig 1 shows that the volume of collected waste has increased significantly over the years, while the weight of collected waste has fluctuated but shows a general downward trend. This suggests that the composition of the waste is changing. Perhaps people are throwing away more bulky items that take up more space but weigh less, or there are efforts underway to reduce denser materials like food scraps or metals.

We create a line plot to visualize the trend of type of the collected waste across different years.

```{r}
# line chart to show the sum of PlasticBottles, Polystyrene, CigaretteButts, GlassBottles, PlasticBags, Wrappers, SportsBalls, HomesPowered by year, log scale for y axis

ggplot(sum_trashwheel,
    aes(x = Year)) + 
    geom_line(aes(x = Year, y = log(sum_PlasticBottles), color = "PlasticBottles")) +
    geom_line(aes(x = Year, y = log(sum_Polystyrene), color = "Polystyrene")) +
    geom_line(aes(x = Year, y = log(sum_CigaretteButts), color = "CigaretteButts")) +
    geom_line(aes(x = Year, y = log(sum_GlassBottles), color = "GlassBottles")) +
    geom_line(aes(x = Year, y = log(sum_PlasticBags), color = "PlasticBags")) +
    geom_line(aes(x = Year, y = log(sum_Wrappers), color = "Wrappers")) +
    geom_line(aes(x = Year, y = log(sum_SportsBalls), color = "SportsBalls")) +
    geom_line(aes(x = Year, y = log(sum_HomesPowered), color = "HomesPowered")) +
    labs(title = "Fig 2. The trend of type of the collected waste across different years", x = "Year", y = "Log Scale by Number of Items") +
    scale_color_manual(values = c("PlasticBottles" = "red", "Polystyrene" = "blue", "CigaretteButts" = "green", "GlassBottles" = "yellow", "PlasticBags" = "purple", "Wrappers" = "orange", "SportsBalls" = "black", "HomesPowered" = "brown")) +
    theme_minimal() + theme(
  plot.title = element_text(color = "blue", face = "bold", hjust = 0.5))
```

### Discussion:

The Figure 2 illustrates that the number of cigarette butts, polystyrene, and plastic bags has decreased significantly, while the trend of other types of waste has increased slightly over the years from 2014 to 2022. This could be attributed to a number of factors, such as increased use of reusable shopping bags or bans on plastic bags. Additionally, we create additional plots such as stacked area plots to compare the composition of collected trash among different Trash Wheels by moths.

```{r}
# show the sum of Weight, Volume for PlasticBottles, Polystyrene, CigaretteButts, GlassBottles, PlasticBags, Wrappers, SportsBalls, HomesPowered by group of ( month)
sum_trashwheel_month_all <- trashwheel %>% 
    group_by(Month) %>% 
    summarise(sum_Weight = sum(Weight), sum_Volume = sum(Volume), sum_PlasticBottles = sum(PlasticBottles), sum_Polystyrene = sum(Polystyrene), sum_CigaretteButts = sum(CigaretteButts), sum_GlassBottles = sum(GlassBottles), sum_PlasticBags = sum(PlasticBags), sum_Wrappers = sum(Wrappers), sum_SportsBalls = sum(SportsBalls), sum_HomesPowered = sum(HomesPowered))

```

The data will arrange by month before we plot.

```{r}
trash_month_all <- sum_trashwheel_month_all %>%
 mutate(Month = factor(Month, levels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December"), labels = 1:12)) %>%
    arrange(Month) 
```

To create a relative frequency bar chart where the waste types are encoded using color, with the month on the y-axis and the percentage of each waste type on the x-axis, we can use the geom_col() function in ggplot2 along with some data manipulation.

```{r}
library(dplyr)
library(ggplot2)

# Calculate relative frequencies
trash_month_all <- trash_month_all %>%
  mutate(total = sum_PlasticBottles + sum_Polystyrene + sum_CigaretteButts + sum_GlassBottles + sum_PlasticBags + sum_Wrappers + sum_SportsBalls) %>%
  mutate(rel_freq_PlasticBottles = sum_PlasticBottles / total,
         rel_freq_Polystyrene = sum_Polystyrene / total,
         rel_freq_CigaretteButts = sum_CigaretteButts / total,
         rel_freq_GlassBottles = sum_GlassBottles / total,
         rel_freq_PlasticBags = sum_PlasticBags / total,
         rel_freq_Wrappers = sum_Wrappers / total,
         rel_freq_SportsBalls = sum_SportsBalls / total)

# Reshape the data to long format and remove NA values from Month variable
trash_long <- trash_month_all %>%
  select(Month, rel_freq_PlasticBottles, rel_freq_Polystyrene, rel_freq_CigaretteButts, 
         rel_freq_GlassBottles, rel_freq_PlasticBags, rel_freq_Wrappers, rel_freq_SportsBalls) %>%
  na.omit() %>%
  pivot_longer(cols = -Month, names_to = "Waste_Type", values_to = "Relative_Frequency")

# Plot the bar chart
ggplot(trash_long, aes(x = Relative_Frequency, y = Month, fill = Waste_Type)) +
  geom_col(position = "stack") +
  scale_x_continuous(labels = scales::percent_format(scale = 1)) +
  labs(x = "Relative Frequency (%)", y = "Month", fill = "Waste Type") +
  ggtitle("Fig 3. Relative Frequency Bar Chart of Waste Types by Month") +
  theme(legend.position = "right") +
  theme(plot.title = element_text(color = "blue", face = "bold", hjust = 0.5))
```

### Discussion:

The Figure 3 illustrates the relationship between the types of collected waste over the months. The frequency of collected cigarette butts was the highest among all types over the months, while sports balls or glass bottles were rarely collected. This suggests that people often discard cigarette butts after using them. Additionally, the frequency of collected cigarette butts increases from May to September, while the frequency of other types of waste increases from January to April. However, the number of items cannot show the trend of weight and volume over the months. Therefore, we will investigate the trend of weight and volume of collected waste over the months in 2018 by using a combination of line chart and bar chart.

```{r}
# show the sum of Weight, Volume for PlasticBottles, Polystyrene, CigaretteButts, GlassBottles, PlasticBags, Wrappers, SportsBalls, HomesPowered by group of (year, month)
sum_trashwheel_month <- trashwheel %>% 
    group_by(Year, Month) %>% 
    summarise(sum_Weight = sum(Weight), sum_Volume = sum(Volume), sum_PlasticBottles = sum(PlasticBottles), sum_Polystyrene = sum(Polystyrene), sum_CigaretteButts = sum(CigaretteButts), sum_GlassBottles = sum(GlassBottles), sum_PlasticBags = sum(PlasticBags), sum_Wrappers = sum(Wrappers), sum_SportsBalls = sum(SportsBalls), sum_HomesPowered = sum(HomesPowered))

trash_month_2018 <- sum_trashwheel_month %>% 
    filter(Year == 2018)
# shorten month name, e.g. January -> Jan and sort by month
trash_month_2018 <- sum_trashwheel_month %>% 
    filter(Year == 2018) %>%
    mutate(Month = factor(Month, levels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December"), labels = 1:12)) %>%
    arrange(Month)


```

```{r}

ggplot(trash_month_2018, aes(x = Month)) +
    geom_bar(aes(y = sum_Weight, fill = "Weight"), stat = "identity", color = "black", alpha = 0.5) +  # Bar for weight
    geom_line(aes(y = sum_Volume / 5, color = "Volume"), group = 1, size = 3) +  # Line for volume
    geom_text(aes(y = sum_Weight, label = sum_Weight), hjust = 0.5, vjust = -0.5, color = "red", size = 4) +
    scale_y_continuous(
        name = "Weight",
        sec.axis = sec_axis(~ . * 5, name = "Volume")
    ) +
    labs(title = "Fig 4. Weight and Volume by Month in 2018", x = "Month", y = "Weight") +
    scale_color_manual(values = c("Volume" = "blue"), labels = c("Volume")) +  # Legend for volume
    scale_fill_manual(values = c("Weight" = "red"), labels = c("Weight")) +  # Legend for weight
    theme(
        axis.text = element_text(size = 15),
        axis.title = element_text(size = 15, face = "bold"),
        plot.title = element_text(size = 20, face = "bold")
    ) +
    guides(fill = guide_legend(title = "Data Type"), color = guide_legend(title = "Data Type"))  # Add legend titles for both fill and color aesthetics
    


```

### Discussion:

The Figure 4 illustrates that the weight and volume of the collected waste from April to June are higher than in other months. Furthermore, the trends of weight and volume remain consistent across the months in 2018.

# Question 2: What is the seasonal pattern of trash accumulation in terms of weight (Weight), volume (Volume), and the number of plastic bags (PlasticBags)?

## Introduction:

The investigation into the seasonal pattern of trash accumulation in terms of weight, volume, and the number of plastic bags arises from a critical need to comprehend the dynamic nature of waste management practices across different temporal scales. Central to this inquiry are several key variables within the dataset, namely Month, Year, Weight, Volume, and the count of PlasticBags. These variables serve as indispensable components for unraveling the intricate relationship between seasonal fluctuations and trash collection trends. By scrutinizing variations in these parameters over different months and years, we aim to uncover discernible patterns that shed light on the underlying factors driving trash accumulation.

Understanding the seasonal dynamics of trash accumulation holds significant implications for environmental stewardship and sustainable waste management strategies. The exploration of weight, volume, and plastic bag counts across various timeframes enables us to discern not only the temporal trends in trash accumulation but also the potential influence of seasonal factors on waste generation and disposal patterns. By delving into these nuances, we aspire to glean insights that can inform targeted interventions and policies aimed at mitigating the environmental impact of waste accumulation, thus fostering a healthier and more sustainable future for our communities and ecosystems. 
## Approach: 
To unravel the seasonal patterns of trash accumulation in terms of weight, volume, and the number of plastic bags, we adopt a systematic approach employing diverse visualizations tailored to capture distinct facets of the data. Initially, we organize the dataset by quarter and year to streamline our analysis. Our first focus lies on delineating the seasonal trajectory of plastic bottle accumulation. Through the utilization of bar charts, we illustrate the temporal trends in collected plastic bottles across the years from 2016 to 2022. This choice of visualization enables us to discern any discernible seasonal fluctuations or patterns in plastic bottle accumulation over time. Subsequently, we extend our analysis to encompass the weight and volume of collected waste. Leveraging additional bar charts, we portray the temporal dynamics of weight and volume over the quarters spanning from 2016 to 2022. These visualizations offer a comprehensive depiction of the overall trends in trash accumulation over the specified timeframe.

To delve deeper into the interrelationships between different types of trash accumulation, we embark on an exploration of potential correlations. Employing a heatmap visualization, we elucidate the correlation matrix between various types of trash, including plastic bottles, polystyrene items, glass bottles, cigarette butts, wrappers, and sports balls, with a particular focus on the year 2015. By employing color mapping to represent the strength of correlations, this visualization aids in identifying significant associations between different types of trash accumulation. Furthermore, we delve into the distribution of weight of collected waste across quarters utilizing a box plot. This visualization offers insights into the variability and central tendency of weight distribution within each quarter, providing valuable insights into the seasonal patterns of trash accumulation in terms of weight. Through a judicious combination of bar charts, heatmap, and box plot, our analytical approach offers a comprehensive understanding of the seasonal dynamics of trash accumulation, shedding light on weight, volume, and the prevalence of plastic bags over time. 
## Analysis: 
Firstly, we want to visualize the seasonal pattern of trash accumulation for Plastic Bottles. Therefore, we create the data.

```{r}
# show the sum of Weight, Volume for PlasticBottles, Polystyrene, CigaretteButts, GlassBottles, PlasticBags, Wrappers, SportsBalls, HomesPowered by group of (year, quarter)
# the month in text "January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December" converted to quarter "Q1", "Q2", "Q3", "Q4"
trashwheel <- trashwheel %>% 
    mutate(quarter = ifelse(Month %in% c("January", "February", "March"), "Q1", ifelse(Month %in% c("April", "May", "June"), "Q2", ifelse(Month %in% c("July", "August", "September"), "Q3", "Q4"))))

sum_trashwheel_quarter <- trashwheel %>% 
    group_by(Year, quarter) %>%
    summarise(sum_Weight = sum(Weight), sum_Volume = sum(Volume), sum_PlasticBottles = sum(PlasticBottles), sum_Polystyrene = sum(Polystyrene), sum_CigaretteButts = sum(CigaretteButts), sum_GlassBottles = sum(GlassBottles), sum_PlasticBags = sum(PlasticBags), sum_Wrappers = sum(Wrappers), sum_SportsBalls = sum(SportsBalls), sum_HomesPowered = sum(HomesPowered))
```

```{r}
# check number of rows by Year

# only select data in year_list
year_list <- c(2015, 2016, 2018, 2020, 2021, 2022)

sec_year_trashwheel <- trashwheel %>% 
    filter(Year %in% year_list)

trashwheel_quater <- sec_year_trashwheel %>% 
    mutate(quarter = ifelse(Month %in% c("January", "February", "March"), "Q1", ifelse(Month %in% c("April", "May", "June"), "Q2", ifelse(Month %in% c("July", "August", "September"), "Q3", "Q4"))))

sum_trashwheel_quarter <- trashwheel_quater %>% 
    group_by(Year, quarter) %>%
    summarise(sum_Weight = sum(Weight), sum_Volume = sum(Volume), sum_PlasticBottles = sum(PlasticBottles), sum_Polystyrene = sum(Polystyrene), sum_CigaretteButts = sum(CigaretteButts), sum_GlassBottles = sum(GlassBottles), sum_PlasticBags = sum(PlasticBags), sum_Wrappers = sum(Wrappers), sum_SportsBalls = sum(SportsBalls), sum_HomesPowered = sum(HomesPowered))

```

Plot the bar chart to show the trend of the collected plastic bottles over the seasons from 2016 to 2022.

```{r}
# show PlasticBottles by year and quarter
ggplot(sum_trashwheel_quarter, aes (x = Year, y = sum_PlasticBottles, fill = quarter)) +
    geom_col(position = "dodge") +
    labs(title = "Fig 5. Plastic Bottles by Year and Quarter", x = "Year", y = "PlasticBottles", color = "Type") +
    theme_minimal() + 
    theme(
        axis.text = element_text(size = ),
        axis.title = element_text(size = 15, face = "bold"),
        plot.title = element_text(size = 20, face = "bold")
    ) +
    coord_flip() 
```

### Discussion:

The Fig 5. illustrates the trend of the number of collected waste by year and quarter. The number of collected waste is highest during the summer months (depicted in green), while it is at its lowest during winter (Q4). Additionally, there is a decrease in the number of collected waste observed over the years from 2016 to 2022. We also plot the bar chart to show the weight and volume of collected waste over the quarter from 2016 to 2022.

```{r}
# show Weight by year and quarter
ggplot(sum_trashwheel_quarter, aes (x = Year, y = sum_Weight, fill = quarter)) +
    geom_col(position = "dodge") +
    labs(title = "Fig 6. Weight by Year and Quarter", x = "Year", y = "Weight (tons)", color = "Type") +
    theme_minimal() + 
    theme(
        axis.text = element_text(size = 15),
        axis.title = element_text(size = 15, face = "bold"),
        plot.title = element_text(size = 20, face = "bold")
    ) +
    coord_flip()
```

### Discussion:

Fig. 6 illustrates that the weight of collected waste is highest in the summer (Q2) and autumn (Q3). Conversely, the weight of collected waste is lowest in the spring. Additionally, there is a significant decrease in the weight of collected waste observed from 2016 to 2022. We also plot the bar chart to show the volume of collected waste over the quarter from 2016 to 2022.

```{r}
# show Volume by year and quarter
ggplot(sum_trashwheel_quarter, aes (x = Year, y = sum_Volume, fill = quarter)) +
    geom_col(position = "dodge") +
    labs(title = "Fig 7. Volume by Year and Quarter", x = "Year", y = "Volume (cubic yards)", color = "Type") +
    theme_minimal() + 
    theme(
        axis.text = element_text(size = 15),
        axis.title = element_text(size = 15, face = "bold"),
        plot.title = element_text(size = 20, face = "bold")
    ) +
    coord_flip()
```

### Discussion:

The Fig 7 shows that the trend of volume of collected waste decreases sharply over the 6 years from 2016 to 2022. The volume of collected waste is highest during the summer and lowest in the spring. Determine if there are significant correlations between different types of trash accumulation. We create a heatmap to visualize the correlation matrix between different types of trash (e.g., plastic bottles, polystyrene items, glass bottles, cigarette butts, wrappers, and sports balls) in 2015.

```{r}
# show the heatmap of different types of trash by year and quarter
data_2015 <- sum_trashwheel_quarter %>% 
  filter(Year == 2015) %>%
  select(-Year, -quarter) %>%
  cor()

# remove NA values in the correlation matrix
data_2015[is.na(data_2015)] <- 0

# Set the size of the main title and axis labels
par(cex.main = 0.8, cex.axis = 0.8)

# Plotting heatmap
heatmap(data_2015, Rowv = NA, Colv = NA, col = cm.colors(256), scale = "none",
        main = "Fig 8. Heatmap of Different Types of Trash by Year and Quarter (2015)")
```

### Discussion:

The Fig 8 shows the correlation between different types of trash accumulation. The correlations between plastic bottles, polystyrene, cigarette butts, and sports balls are very strong, whereas the correlations between other collected waste are weak. Furthermore, strong correlations are observed between weight, volume, plastic bottles, polystyrene, cigarette butts, and homes powered. This suggests that these types of collected waste contribute significantly to power generation, and an increase in weight and volume will result in an increase in homes powered. We also investigate the distribution of weight of collected waste over the quarter by using the box plot.

```{r}
# box plot to show the distribution of Weight by quarter
ggplot(trashwheel_quater, aes(x = quarter, y = Weight, fill = quarter)) +
    geom_boxplot() +
    labs(title = "Fig 9. Distribution of Weight by Quarter", x = "Quarter", y = "Weight (tons)", color = "Type") +
    theme_minimal() + 
    theme(
        axis.text = element_text(size = 15),
        axis.title = element_text(size = 15, face = "bold"),
        plot.title = element_text(size = 20, face = "bold")
    )
```

### Discussion:

The weight of collected waste is highest in the first quarter (Q1) with a median of around 4 tons. The weight is lowest in the second quarter (Q2) with a median of around 1 ton. There seems to be more variability in the weight collected in Q3 compared to Q1 and Q4. The median for Q3 is at around 2 tons, however the whiskers extend further out than the other quarters which signifies a wider range of data points. Overall, the spread of the data in Q1 is smaller than Q3 and Q4, which means the weights in Q1 are more clustered around the median.

# Conclusion:

In conclusion, our analysis has shed light on two key aspects of trash collection: the variation in composition across different Trash Wheels over time, and the seasonal patterns of trash accumulation in terms of weight, volume, and the number of plastic bags.

For the first question, we utilized a variety of visualizations including line charts, bar charts,dual Y-axis bar and line charts, and frequency bar charts to explore the relationship between the number of plastic bottles, polystyrene items, and glass bottles collected by each Trash Wheel across different months and years. Our analysis revealed significant variations in trash composition among different Trash Wheels, providing valuable insights into the dynamics of waste accumulation over time.

Regarding the second question, we employed bar charts, heatmaps, and box plots to investigate the seasonal patterns of trash accumulation. Our findings highlighted distinct seasonal trends, with fluctuations in weight, volume, and the number of plastic bags observed across different months and years. Additionally, the heatmap analysis revealed correlations between different types of waste, offering further insights into waste management strategies.

Overall, our comprehensive analysis has enhanced our understanding of trash collection dynamics, providing valuable insights that can inform waste management policies and practices aimed at optimizing collection efficiency and environmental sustainability.
